Temporal graph realization from fastest paths

### 2024-01-17 SAND2024/SAND24-longversion.tex
Finished with the changes for the FPT proof.

    \begin{definition}[Travel delays]
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\subsubsection{Adding constraints and variables to the ILP}\label{sec:constraints}
We start by analyzing the case where we want to determine the labels on fastest temporal paths between vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-LabelAlmostalledges}).
We call these $\ell - 1$ edges, \emph{partially determined} edges.
After repeating this step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge of each path,
or are partially determined.
If the fastest temporal path between two vertices $u,v \in U$ is just an edge $e$, then we treat it as being determined, since it gets assigned a label $\lambda(e)$ with respect to itself.
All other edges in $G'$ %that were not determined with the respect to some other edge label up to now, or are not partially determined, 
are called the \emph{not yet determined} edges.
Note that the not yet determined edges are exactly the ones that are not a part of any fastest temporal path between any two vertices in $U$.

@@

From the above procedure (where we were determining labels of edges of segments with each other) 
we conclude that all of the edges $e_i=v_i v_{i+1}$ of a segment $S_{u,v}=(u=v_1, v_2, \dots, v_p=v)$ are in one of the following relations.
First, where all of the edges are determined with respect to each other.
Second, where
there are some edges $e_1, e_2, \dots e_{i-1}$, whose label is determined with respect to the label $\lambda(e_1)$, there is an edge $f = e_i = v_i v_{i+1}$ which is not yet determined,
and then there follow the edges
$e_{i+1}, e_{i+2}, \dots, e_{p-1}$, whose labels are determined with respect to $\lambda(e_{p-1})$.
Third, where the first $e_1, \dots, e_{i-1}$ edges are determined with respect to the $\lambda(e_1)$ and all of the remaining edges $e_i, e_{i+1} \dots, e_{p-1}$ are determined with respect to the $\lambda(e_{p-1})$.
%
We want to now determine all of the edges in such segment $S_{u,v}$ with respect to just one edge (either the first or the last one).
In the second case, we use the fact that at least one of the temporal paths between $v_{i-1}$ and $v_{i+1}$ has to pass through $f$, to determine $\lambda(f)$ with respect to $\lambda(e_{i-1})$ (and consequently $\lambda(e_1)$),
and similarly, one of the temporal paths between $v_{i}$ and $v_{i+2}$ has to pass through $f$, which determines $\lambda(f)$ with respect to $\lambda(e_{i+1})$ (and consequently $\lambda(e_{p-1})$).
In the third case, knowing the temporal paths between $v_{i-1}$ to $v_{i+1}$ results in determining the label of $\lambda(e_{i-1}$ with $\lambda(e_i)$, which consequently relates labels of all of the edges of the segment against each other.
%
To determine the desired paths we proceed as follows.
\begin{enumerate}[G-1.]
    \setcounter{enumi}{\value{guesscounter}}
    \item \label{FPT-guessFTPinSegmentTgroughEdge}
    Let $S_{u,v} = (u=v_1, v_2, \dots, v_p=v)$  be a segment of length at least $4$.
    If there is a not yet determined edge $v_i v_{i+1} = f$ in $S_{u,v}$ then
    we guess which of the fastest temporal paths: from $v_{i-1}$ to $v_{i+1}$, from $v_{i+1}$ to $v_{i-1}$, from $v_{i}$ to $v_{i+2}$, from $v_{i+2}$ to $v_{i}$ pass through the edge $f$. 
    If there are two incident edges $e = v_{i-1} v_{i}$ and $f = v_i v_{i+1} $ in $S_{u,v}$, that are determined with respect to $\lambda(v_1 v_2)$ and $\lambda (v_{p-1} v_p)$, respectively then we guess which of the fastest temporal paths: from $v_{i-1}$ to $v_{i+1}$, from $v_{i+1}$ to $v_{i-1}$ pass through the edges $e,f$. \\
    %
    We create $O(1)$ guesses for every such segment $S_{u,v}$, and $O(k^2)$ new guesses in total, as there are at most $O(k^2)$ segments.
    % counter for guesses continues
    \setcounter{guesscounter}{\value{enumi}}
\end{enumerate}
%
Note that the condition for segment length at least four comes from~\cref{lemma:FPT-noUndeterminedEdgeInSegment}.
We now conclude the following.
\begin{observation}\label{obs:FPT-4determined}
    Let $S_{u,v}$ be an arbitrary segment in $G'$.
    If $S_{u,v}$ is of length at least $4$ then the labels of all the edges are determined with respect to its first edge.
    If $S_{u,v}$ is of length $3$ or $2$ then it has at most $3$ or $2$ not yet determined edges, respectively.
\end{observation}

@@

\subparagraph{\boldmath Fastest paths from $x \in V(G') \setminus U$ to $u \in U$.}
% Now we proceed with the case of %determining the fastest path from $x$ to $u$, where $x \in V(G') \setminus U$ and $u \in U$.
%We use the similar approach as in the case~\ref{item:FPT-uToxcase2} from above.
Let $x$ be a vertex in the segment $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$, and let $u\in U$.
If $S_{w,z}$ is of length $3$ or less, then we already know the fastest temporal path from every vertex in the segment to $u$ 
(since $S_{w,z}$ has at most $2$ inner vertices, we determined the fastest temporal paths from them to $u$ in guess \textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongUandUstar}}}}).
Assume that $S_{w,z}$ is of length at least $4$.

This implies that there are at most two not yet determined edges in it.
We can easily compute the vertices $z_i, z_{i+1} \in S_{w,z}\setminus\{w,z\}$ for which the fastest temporal path from $z_i$ to $u$ has the biggest duration.
Denote with $P^+$ the fastest temporal path of the form $z_2 \rightarrow z \leadsto u$,
and with $P^-$ the fastest temporal path of the form $z_{r-1} \rightarrow w \leadsto u$.
Note that we know these paths from guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guess-uToSegmentz2}}}}.
Then we know that all vertices $z_j$ in $S_{w,z} \setminus\{z_i,z_{i+1}\}$ that are closer to $w$ than $z_i,z_{i+1}$ reach $u$ on the 
following fastest temporal path $(z_j \rightarrow z_{j-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$ 
and
all the vertices $z_j$ in $S_{w,z}\setminus\{z_i,z_{i+1}\}$ that are closer to $z$ than $z_i, z_{i+1}$ reach $u$ on the 
following fastest temporal path
$(z_j \rightarrow z_{j+1} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
Since the first part of fastest paths is unique, and we know the second part is the fastest, the above paths are the fastest temporal paths.
We still have to determine the fastest temporal paths from $z_i,z_{i+1}$ to $u$.
We distinguish the following two options.
\begin{enumerate}[(i)]
    \item \label{FPT:equality-fromUtoX-twosplit}
    $z_i \neq z_{i+1}$.
    Then the fastest temporal path from $z_i$ to $u$ is 
    $(z_i \rightarrow z_{i-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$,
    and 
    the fastest temporal path from $z_{i+1}$ to $u$ is 
    $(z_{i+1} \rightarrow z_{i+2} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
    \item \label{FPT:equality-fromUtoX-onesplit}
    $z_i = z_{i+1}$, \ie let $z_i$ be the unique vertex, that is furthest away from $u$ in $S_{w,z}$.
    In this case we have to determine if the fastest temporal path from $z_i$ to $u$, goes first through vertex $z_{i-1}$ (and then through $w$),
    or it goes first through $z_{i+1}$ (and then through $z$).
    Since we know the values $D_{z_{i-1},u}, D_{z_{i+1},u}$,
    and since there are at most two not yet determined labels in $S_{w,z}$,
    we can uniquely determine one of the following
    waiting times:
    the waiting time $\tau_{v_{i-1}} ^{v_{i}, v_{i-2}}$ at vertex $v_{i-1}$ when traveling from $v_i$ to $v_{i-2}$,
    or
    the waiting time $\tau_{v_{i+1}} ^{v_{i}, v_{i+2}}$ at vertex $v_{i+1}$ when traveling from $v_i$ to $v_{i+2}$.
    Suppose we know the former (for the latter case, analysis follows similarly).
    Then we set $c = D_{z_{i-1},u} +  \tau_{v_{i-1}} ^{v_{i}, v_{i-2}}$.
    We now compare $c$ and the value $D_{z_{i},u}$.
    If $c < D_{z_{i},u}$ we conclude that our ILP has no solution and we stop with calculations,
    if $c = D_{z_{i},u}$ then the fastest temporal path from $z_i$ to $u$ is of the form $(z_i \rightarrow z_{i-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$,
    if $c > D_{z_{i},u}$ then the fastest temporal path from $z_i$ to $u$ is of the form $(z_i \rightarrow z_{i+1} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
\end{enumerate}
Once the fastest temporal path from $c$ to $u$ is determined, we introduce an equality constraint for it.
For each of the other $k^{O(k)}$ paths from $x$ to $u$ (which correspond to all paths of the form $w \leadsto u$ and $z \leadsto u$, together with the unique subpath on $S_{w,z}$),
we introduce an inequality constraint. Therefore we add $k^{O(k)}$ inequality constraints for the pair $x,u$.

\subparagraph{\boldmath Fastest paths between $x,y\in V(G') \setminus U$.}
Let $x,y \in V(G') \setminus U$.
We have two options.
\begin{enumerate}[(i)]
    \item Vertices $x,y$ are in the same segment $S_{u,v} = (u,v_1,v_2, \dots, v_p, v)$. 
    If the length of $S_{u,v}$ is less than $4$ then we know what is the fastest path between vertices.
    Suppose now that $S_{u,v}$ is of length at least $5$. Then there are at most two not yet determined edges in $S_{u,v}$.

    W.l.o.g. suppose that $x$ is closer to $u$ in $S_{u,v}$ than $y$.
    Denote with $x = v_i$.
    Let $v_k \in S_{u,v}$ be a vertex for which the duration from $v_i$ is the biggest 
    (note that in the case when we have two such vertices, $v_k$ and $v_{k+1}$ we know exactly what are the fastest paths from $x$ to every vertex in $S_{u,v}$,
    by similar arguing as in case~(\ref{FPT:equality-fromUtoX-twosplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$.
    Then we know that $D_{x,v_{i+1}} < D_{x,v_{i+2}} < \cdots  < D_{x,v_{k}}$ and $D_{x,v_{i-1}} < D_{x,v_{i-2}} < \cdots  < D_{x,v_{k}}$, where indices are taken modulo $p$.
    Therefore we know exactly the structure of all the fastest paths from $x$ to every vertex in $S_{u,v}$,
    with the exception of the fastest path from $x$ to $v_k$.
    Since there is at most one undetermined edge in $S_{u,v}$,
    and since we know the exact durations $D_{x,v_{k-1}}$ and $D_{x,v_{k+1}}$,
    we can determine either
    $c = D_{x,v_{k-1}} + \tau _ {v_k-1}^{v_k,v_{k-2}}$ or 
    $c' = D_{x,v_{k+1}} + \tau _ {v_k+1}^{v_k,v_{k+2}}$.
    We then compare (one of) these values to $D_{x,v_k}$ which then 
    uniquely determines the fastest temporal path from $x$ to $v_k$
    (for details see case~(\ref{FPT:equality-fromUtoX-onesplit}) from above, when we were determining the fastest path from $x \in V(G')\setminus U$ to $u \in U$).
    \item Vertices $x$ and $y$ are in different segments. 
Let $x$ be a vertex in the segment $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and let $y$ be a vertex in the segment $S_{w,z} = (w=z_1, z_2,z_3, \dots, z_r = z)$.
By checking the durations of the fastest paths from $x$ to every vertex in $S_{w,z} \setminus \{w,z\}$
we can determine the vertex $z_i \in S_{w,z} $, for which the duration from $x$ is the biggest.
Note that if there are two such vertices $z_i$ and $z_{i+1}$, we know exactly how all fastest temporal paths enter $S_{w,z}$ 
(we use similar arguing as in case~(\ref{FPT:equality-fromUtoX-twosplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$).
This implies that the fastest temporal paths from $x$ to all vertices $z_2, z_3, \dots, z_{i-1}$ (resp.~$z_{i+1}, z_{i+2}, \dots, z_{r-1}$)  pass through $w$ (resp.~$z$).
Now we determine the vertex $v_j \in S_{u,v} \setminus \{u,v\}$,
for which the value of the durations of the fastest paths from it to the vertex $y$ is the biggest.
Again, if there are two such vertices $v_{j}$ and $v_{j+1}$ we know exactly how the fastest temporal paths, starting in these two vertices,
leave the segment $S_{u,v}$. 
We use similar arguing as in case~(\ref{FPT:equality-fromUtoX-twosplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$.
Knowing the vertex $v_j$
implies that the fastest temporal paths from the vertices $v_2, v_2, \dots, v_{j-1}$ (resp.~$v_{j+1}, v_{j+2}, \dots, v_{p-1}$) to the vertex $y$ passes through $u$ (resp.~$v$).
Since we know the following fastest temporal paths (see guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}}) 
$z_2 \rightarrow w \leadsto u \rightarrow v_2$,
$z_2 \rightarrow w \leadsto v \rightarrow v_{p-1}$,
$z_{r-1} \rightarrow z \leadsto v \rightarrow v_{p-1}$ and
$z_{r-1} \rightarrow z \leadsto v \rightarrow v_{p-1}$,
we can uniquely determine all fastest temporal paths from $x \neq v_j$ to any $y \in S_{u,v} \setminus \{z_i\}$.

We have to now consider the case when 
$x = v_j$ and $y = z_i$.
If at least one of the segments $S_{u,v}$ and $S_{w,z}$ is of length more than $5$,
then this segment has no inner edges with not yet determined labels
and
we can uniquely determine the fastest path from~$v_j$ to $z_i$,
using 
similar arguing as in case~(\ref{FPT:equality-fromUtoX-onesplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$.
If at least one of them is of length $3$ or less,
we can again uniquely determine the fastest path from~$v_j$ to~$z_i$, using the same approach, 
and the knowledge of fastest paths to (or from) all vertices of the segment of length~$3$
(as we guessed them in guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}}).
If both segments are of the length $4$, then we know how all vertices reach each other,
as we guessed the fastest paths in guesses~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}} and~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guess-splitFromAnotherSegmentAndPaths}}}}.
\end{enumerate}
Once the fastest path is determined we introduce the equality constraint for it
and iterate through all other paths, for which we introduce inequality constraints.
To enumerate all these non-fastest temporal paths, we just consider all possible paths $u \leadsto w$, where 
$u$ and $w$ are the vertices of interest that are the endpoints of segments to which $x$ and $y$ belong;
once the correct segment is reached, there is a unique path to the desired vertex $x$ (resp.~$y$). 
Therefore we introduce $k^{O(k)}$ inequality constraints for each pair of vertices $x,y$.

@@

%%% TODO NKL
Recall that $G[Z]$ consists of disjoint trees and that each of these trees has a unique neighbor (clip vertex) $v$ in $G'$. We then define the tree $T_v$ in $G[Z \cup \{v\}]$ as the collection of trees from $G[Z]$ with a clip vertex $v$, together with the root $v$. 
Determining the fastest temporal paths between any two vertices in the same tree is a straightforward process (see~\ref{thm:deltaExact-PolyTimeTrees}), so we exclude this case from our upcoming analysis.
From \cref{obs:tree-only-neighbors} it follows that knowing temporal paths between any $y  \in V(G') \setminus (T_v)$ and all vertices in the first layer of the tree $T_v$ (\ie children of the root $v$),
it is enough to determine the fastest temporal paths between $y$ and all other vertices in $T_v$.
Therefore,
in the upcoming analysis, we focus only on the vertices in the first layer of each tree $T_v$.
During the process of determining the fastest paths from and to the vertices in $Z$, we use the fact that we have already identified the fastest paths among all vertices in $G'$.

>>>>

    \begin{definition}[Travel delays] \label{def:travel-delay}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\subsubsection{Adding constraints and variables to the ILP}\label{sec:constraints}
We start by analyzing the case where we want to determine the labels on fastest temporal paths between vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-Labelalledges}).
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-LabelAlmostalledges}).
We call these $\ell - 1$ edges, \emph{partially determined} edges.
After repeating this step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge of each path,
or are partially determined.
If the fastest temporal path between two vertices $u,v \in U$ is just an edge $e$, then we treat it as being determined, since it gets assigned a label $\lambda(e)$ with respect to itself.
All other edges in $G'$ %that were not determined with the respect to some other edge label up to now, or are not partially determined, 
are called the \emph{not yet determined} edges.
Note that the not yet determined edges are exactly the ones that are not a part of any fastest temporal path between any two vertices in $U$.

@@

From the above procedure (where we were determining labels of edges of segments with each other) 
we conclude that all of the edges $e_i=v_i v_{i+1}$ of a segment $S_{u,v}=(u=v_1, v_2, \dots, v_p=v)$ are in one of the following relations.
First, where all of the edges are determined with respect to each other.
Second, where
there are some edges $e_1, e_2, \dots e_{i-1}$, whose label is determined with respect to the label $\lambda(e_1)$, there is an edge $f = e_i = v_i v_{i+1}$ which is not yet determined,
and then there follow the edges
$e_{i+1}, e_{i+2}, \dots, e_{p-1}$, whose labels are determined with respect to $\lambda(e_{p-1})$.
Third, where the first $e_1, \dots, e_{i-1}$ edges are determined with respect to the $\lambda(e_1)$ and all of the remaining edges $e_i, e_{i+1} \dots, e_{p-1}$ are determined with respect to the $\lambda(e_{p-1})$.
%
We want to now determine all of the edges in such segment $S_{u,v}$ with respect to just one edge (either the first or the last one).
In the second case, we use the fact that at least one of the temporal paths between $v_{i-1}$ and $v_{i+1}$ has to pass through $f$, to determine $\lambda(f)$ with respect to $\lambda(e_{i-1})$ (and consequently $\lambda(e_1)$),
and similarly, one of the temporal paths between $v_{i}$ and $v_{i+2}$ has to pass through $f$, which determines $\lambda(f)$ with respect to $\lambda(e_{i+1})$ (and consequently $\lambda(e_{p-1})$).
In the third case, knowing the temporal paths between $v_{i-1}$ to $v_{i+1}$ results in determining the label of $\lambda(e_{i-1}$ with $\lambda(e_i)$, which consequently relates labels of all of the edges of the segment against each other.
%
To determine the desired paths we proceed as follows.
\begin{enumerate}[G-1.]
    \setcounter{enumi}{\value{guesscounter}}
    \item \label{FPT-guessFTPinSegmentTgroughEdge}
    Let $S_{u,v} = (u=v_1, v_2, \dots, v_p=v)$  be a segment of length at least $4$.
    If there is a not yet determined edge $v_i v_{i+1} = f$ in $S_{u,v}$ then
    we guess which of the fastest temporal paths: from $v_{i-1}$ to $v_{i+1}$, from $v_{i+1}$ to $v_{i-1}$, from $v_{i}$ to $v_{i+2}$, from $v_{i+2}$ to $v_{i}$ pass through the edge $f$. 
    If there are two incident edges $e = v_{i-1} v_{i}$ and $f = v_i v_{i+1} $ in $S_{u,v}$, that are determined with respect to $\lambda(v_1 v_2)$ and $\lambda (v_{p-1} v_p)$, respectively then we guess which of the fastest temporal paths: from $v_{i-1}$ to $v_{i+1}$, from $v_{i+1}$ to $v_{i-1}$ pass through the edges $e,f$. \\
    %
    We create $O(1)$ guesses for every such segment $S_{u,v}$, and $O(k^2)$ new guesses in total, as there are at most $O(k^2)$ segments.
    % counter for guesses continues
    \setcounter{guesscounter}{\value{enumi}}
\end{enumerate}
%
Note that the condition for segment length at least four comes from~\cref{lemma:FPT-noUndeterminedEdgeInSegment}.
We now conclude the following.
\begin{corollary}\label{cor:FPT-4determined}
    Let $S_{u,v}$ be an arbitrary segment in $G'$.
    If $S_{u,v}$ is of length $3$ or $2$ then it has at most $3$ or $2$ not yet determined edges, respectively.
    If $S_{u,v}$ is of length at least $4$ then the labels of all its edges are determined with respect to the first edge. 
\end{corollary}

@@

\subparagraph{\boldmath Fastest paths from $x \in V(G') \setminus U$ to $u \in U$.}
% Now we proceed with the case of %determining the fastest path from $x$ to $u$, where $x \in V(G') \setminus U$ and $u \in U$.
%We use the similar approach as in the case~\ref{item:FPT-uToxcase2} from above.
Let $x$ be a vertex in the segment $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$, and let $u\in U$.
If $S_{w,z}$ is of length $3$ or less, then we already know the fastest temporal path from every vertex in the segment to $u$ 
(since $S_{w,z}$ has at most $2$ inner vertices, we determined the fastest temporal paths from them to $u$ in guess \textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongUandUstar}}}}).

Assume that $S_{w,z}$ is of length at least $4$. From \cref{cor:FPT-4determined} we know that the labels of all the edges in $S_{w,z}$ are determined with respect to the label of the first edge.
Moreover, this gives us the knowledge of the exact differences among two consecutive edge labels, which is enough to uniquely determine travel delays at all of the inner vertices $z_i \in S_{w,z}$ (see~\cref{def:travel-delay}).

From the matrix $D$ we can easily determine the two vertices $z_i, z_{i+1} \in S_{w,z}\setminus\{w,z\}$ for which the fastest temporal path from $z_i$ to $u$ has the biggest duration.
Let us denote with $P^+$ the fastest temporal path of the form $z_2 \rightarrow z \leadsto u$,
and with $P^-$ the fastest temporal path of the form $z_{r-1} \rightarrow w \leadsto u$ 
(we know these paths from guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guess-uToSegmentz2}}}}).
%
It follows that all vertices $z_j$ in $S_{w,z} \setminus\{z_i,z_{i+1}\}$ that are closer to $w$ than $z_i,z_{i+1}$ reach $u$ the fastest using the path $(z_j \rightarrow z_{j-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$ 
and
all the vertices $z_j$ in $S_{w,z}\setminus\{z_i,z_{i+1}\}$ that are closer to $z$ than $z_i, z_{i+1}$ reach $u$ the fastest using the path
$(z_j \rightarrow z_{j+1} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
Since the first part of the above path is unique, and we know that the second part is the fastest, it follows that these paths indeed represent the fastest temporal paths to $u$.
%
What remains to determine is the fastest temporal paths from $z_i,z_{i+1}$ to $u$.
We distinguish the following two options.
\begin{enumerate}[(i)]
    \item \label{FPT:equality-fromUtoX-twosplit}
    $z_i \neq z_{i+1}$.
    Then the fastest temporal path from $z_i$ to $u$ is 
    $(z_i \rightarrow z_{i-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$,
    and 
    the fastest temporal path from $z_{i+1}$ to $u$ is 
    $(z_{i+1} \rightarrow z_{i+2} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
    \item \label{FPT:equality-fromUtoX-onesplit}
    $z_i = z_{i+1}$, \ie let $z_i$ be the unique vertex, that is furthest away from $u$ in $S_{w,z}$.
    In this case we have to determine if the fastest temporal path from $z_i$ to $u$, travels first through vertex $z_{i-1}$ (and then through $w$),
    or it travels first through $z_{i+1}$ (and then through $z$).
    Since we know the values $D_{z_{i-1},u}, D_{z_{i+1},u}$,
    and we know the value of the waiting time $\tau_{v_{i-1}} ^{v_{i}, v_{i-2}}$ at vertex $v_{i-1}$ when traveling from $v_i$ to $v_{i-2}$, 
    we can uniquely determine the desired path.
    We set $c = D_{z_{i-1},u} +  \tau_{v_{i-1}} ^{v_{i}, v_{i-2}}$
    and compare $c$ to the value $D_{z_{i},u}$.
    If $c < D_{z_{i},u}$ we conclude that our ILP has no solution and we stop with calculations,
    if $c = D_{z_{i},u}$ then the fastest temporal path from $z_i$ to $u$ is of the form $(z_i \rightarrow z_{i-1} \rightarrow \cdots \rightarrow z_2) \cup P^+$,
    if $c > D_{z_{i},u}$ then the fastest temporal path from $z_i$ to $u$ is of the form $(z_i \rightarrow z_{i+1} \rightarrow \cdots \rightarrow z_{r-1}) \cup P^-$.
\end{enumerate}
Once the fastest temporal path from $c$ to $u$ is determined, we introduce an equality constraint for it.
For each of the other $k^{O(k)}$ paths from $x$ to $u$ (which correspond to all paths of the form $w \leadsto u$ and $z \leadsto u$, together with the unique subpath on $S_{w,z}$),
we introduce an inequality constraint. Therefore we add $k^{O(k)}$ inequality constraints for the pair $x,u$.

\subparagraph{\boldmath Fastest paths between $x,y\in V(G') \setminus U$.}
Let $x,y \in V(G') \setminus U$.
There are two options.
\begin{enumerate}[(i)]
    \item Vertices $x,y$ are in the same segment $S_{u,v} = (u,v_1,v_2, \dots, v_p, v)$. 
    If the length of $S_{u,v}$ is less than $4$ then we know what is the fastest path between vertices, as $x,y \in U^*$.
    Suppose now that $S_{u,v}$ is of length at least $4$ and assume that $x$ is closer to $u$ in $S_{u,v}$ than $y$.
    Then we have two options; either the path from $x$ to $y$ travels only through the edges of $S_{u,v}$, denote such path as $P_{x,y}$,
    or it is of the form $x \rightarrow v_1 \rightarrow u \leadsto v \rightarrow v_p \rightarrow y$, denote is as $P_{x,y}^*$.
    Note that we can determined $P_{x,y}^*$ as it is a concatenation of a unique path from $x$ to $v_2$, together with the fastest path from $v_2$ to $v_{p}$, that travels through $u$ and $v$ (we know this path from \textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}}),
    and the unique path from $v_p$ to $y$.
    %
    Because of \cref{cor:FPT-4determined} we can determine $c = d(P_{x,y}, \lambda)$.
    If $c > D_{x,y}$ we set the fastest path to be $P_{x,y}^*$,
    if $c = D_{x,y}$ then the fastest path is $P_{x,y}$, and if 
    $c < D_{x,y}$ we conclude that our ILP has no solution and we stop with calculations.
    %
    \item Vertices $x$ and $y$ are in different segments. 
Let $x$ be a vertex in the segment $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and let $y$ be a vertex in the segment $S_{w,z} = (w=z_1, z_2,z_3, \dots, z_r = z)$.
%
By checking the durations of the fastest paths from $x$ to every vertex in $S_{w,z} \setminus \{w,z\}$
we can determine the vertex $z_i \in S_{w,z} $, for which the duration from $x$ is the biggest.
Note that if there are two such vertices $z_i$ and $z_{i+1}$, we know exactly how all fastest temporal paths enter $S_{w,z}$ 
(we use similar arguing as in case~(\ref{FPT:equality-fromUtoX-twosplit}) from above, where we were determining the fastest path from $x \in V(G')$ to $u \in U$).
This implies that the fastest temporal paths from $x$ to all vertices $z_2, z_3, \dots, z_{i-1}$ (resp.~$z_{i+1}, z_{i+2}, \dots, z_{r-1}$)  pass through $w$ (resp.~$z$).
Now we determine the vertex $v_j \in S_{u,v} \setminus \{u,v\}$,
for which the value of the durations of the fastest paths from it to the vertex $y$ is the biggest.
Again, if there are two such vertices $v_{j}$ and $v_{j+1}$ we know exactly how the fastest temporal paths, starting in these two vertices,
leave the segment $S_{u,v}$. 
We use similar arguing as in case~(\ref{FPT:equality-fromUtoX-twosplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$.
Knowing the vertex $v_j$
implies that the fastest temporal paths from the vertices $v_2, v_2, \dots, v_{j-1}$ (resp.~$v_{j+1}, v_{j+2}, \dots, v_{p-1}$) to the vertex $y$ passes through $u$ (resp.~$v$).
Since we know the following fastest temporal paths (see guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}}) 
$z_2 \rightarrow w \leadsto u \rightarrow v_2$,
$z_2 \rightarrow w \leadsto v \rightarrow v_{p-1}$,
$z_{r-1} \rightarrow z \leadsto v \rightarrow v_{p-1}$ and
$z_{r-1} \rightarrow z \leadsto v \rightarrow v_{p-1}$,
we can uniquely determine all fastest temporal paths from $x \neq v_j$ to any $y \in S_{u,v} \setminus \{z_i\}$.

In case when $x = v_j$ and $y = z_i$ and the segments are of length at least $4$, we can uniquely determine the fastest path from~$v_j$ to $z_i$,
using 
similar arguing as in case~(\ref{FPT:equality-fromUtoX-onesplit}) from above, when we were determining the fastest path from $x \in V(G')$ to $u \in U$.
If at least one of the segments is of length $3$ or less,
we can again uniquely determine the fastest path from~$v_j$ to~$z_i$, using the same approach, 
and the knowledge of fastest paths to (or from) all vertices of the segment of length~$3$
(as we guessed them in guess~\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT-guessFTPamongv2z2}}}}).
\end{enumerate}
Once the fastest path is determined we introduce the equality constraint for it
and iterate through all other paths, for which we introduce inequality constraints.
To enumerate all these non-fastest temporal paths, we just consider all possible paths $u \leadsto w$, where 
$u$ and $w$ are the vertices of interest that are the endpoints of segments to which $x$ and $y$ belong;
once the correct segment is reached, there is a unique path to the desired vertex $x$ (resp.~$y$). 
Therefore we introduce $k^{O(k)}$ inequality constraints for each pair of vertices $x,y$.

@@

Recall that $G[Z]$ consists of disjoint trees and that each of these trees has a unique neighbor (clip vertex) $v$ in $G'$. We then define the tree $T_v$ in $G[Z \cup \{v\}]$ as the collection of trees from $G[Z]$ with a clip vertex $v$, together with the root $v$. 
Determining the fastest temporal paths between any two vertices in the same tree is a straightforward process (see~\ref{thm:deltaExact-PolyTimeTrees}), so we exclude this case from our upcoming analysis.
From \cref{obs:tree-only-neighbors} it follows that knowing temporal paths between any $y  \in V(G') \setminus (T_v)$ and all vertices in the first layer of the tree $T_v$ (\ie children of the root $v$),
it is enough to determine the fastest temporal paths between $y$ and all other vertices in $T_v$.
Therefore,
in the upcoming analysis, we focus only on the vertices in the first layer of each tree $T_v$.
During the process of determining the fastest paths from and to the vertices in $Z$, we use the fact that we have already identified the fastest paths among all vertices in $G'$.
